Psychological Needs During Intergroup Contact: Three Experience Sampling Studies

### 2022-03-14 manuscript.tex
Updates from Overleaf

% define document type (i.e., template. Here: A4 APA manuscript with 12pt font)
\documentclass[man, 12pt, a4paper]{apa7}

@@

% Title and header
\title{Psychological Needs During Intergroup Contact}
\shorttitle{Needs in Intergroup Contact}

% Authors
\author[*,1,2]{Jannis Kreienkamp}
\author[1,2]{Maximilian Agostini}
\author[1,2]{Laura F. Bringmann}
\author[1,2]{Peter de Jonge}
\author[1,2]{Kai Epstude}
\affiliation{\hfill}

\affil[1]{University of Groningen, Department of Psychology}
\affil[2]{Any additional affiliations still need to be added}

@@

We have no known conflict of interest to declare. The authors received no specific funding for this work. Source data and  software is available at \url{https://janniscodes.github.io/intergroup-contact-needs/}. Protocols, materials, analysis data, and code are available at \url{osf.io}.

@@

% Abstract
\abstract{
Modern intergroup contact research faces two main challenges: (1) When does a positively perceived interaction with an outgroup member translate to positive intergroup contact. (2) Do the often between-person studied effects hold within-person? Here we present three extensive longitudinal studies of recent migrants to capture their interactions with the majority outgroup (total \textit{N} of measurements = 10,297). We find that  fulfillment of psychological needs, going through perceived interaction quality, are a strong predictor of positive intergroup attitudes. Across the three studies, we also find inconsistent evidence for interaction frequencies and average interaction quality in predicting outgroup attitudes. Finally, situational needs remained the core predictor of outgroup attitudes even after controlling for Allport's contact conditions or more general fundamental needs. As one of the first studies to test intergroup contact theory using extensive longitudinal data, we offer insight into the mechanisms of positive intergroup contact during real-life interactions and find situational motivations to be a key building block of understanding and addressing positive intergroup interactions.

\noindent\textbf{Public significance statement}: In this paper we provide evidence that the fulfillment of situational psychological needs during real-life intergroup contacts meaningfully predicts perceived interaction quality and positive outgroup attitudes. This offers testament to the emerging practice of capturing real-life interactions using extensive longitudinal data. It also gives weight to motivational fulfillment as a flexible and relevant mechanism in understanding positive intergroup contact.
}

@@

\begin{figure}
  \caption{Contact Hypothesis}
  \label{fig:ContactHypothesis}
  \begin{subfigure}{\textwidth}
    \caption{}
    \centering\includegraphics[width=0.75\textwidth]{Figures/forestParametricGeneralLm.png}
  \end{subfigure}
  \begin{subfigure}{\textwidth}
    \caption{}
    \centering\includegraphics[width=0.75\textwidth]{Figures/forestParametricFEGeneralLmer.png}
  \end{subfigure}
  \caption*{Note: \\
  (a) summary of regression results from the aggregated contact and interaction quality data.\\
  (b) summary of mixed models results of the contemporaneous contact effects.\\
  General: Fixed effects meta analytic results are presented for completeness only.}
\end{figure}

\begin{figure}
  \caption{Core Need Fulfillment}
  \label{fig:AllportNeedFulfillment}
  \begin{subfigure}{\textwidth}
    \caption{}
    \centering\includegraphics[width=0.6\textwidth]{Figures/forestParametricFETheoryQualityCore.png}
  \end{subfigure}
  \begin{subfigure}{\textwidth}
    \caption{}
    \centering\includegraphics[width=0.6\textwidth]{Figures/forestParametricFETheoryAttitudeCore.png}
  \end{subfigure}
  \begin{subfigure}{\textwidth}
    \caption{}
    \centering\includegraphics[width=0.6\textwidth]{Figures/forestParametricFETheoryAttitudeCoreQuality.png}
  \end{subfigure}
  \caption*{Note: \\
  (a) Core Need Fulfillment predicting Interaction Quality.\\
  (b) Core Need Fulfillment predicting Outgroup Attitudes.\\
  (c) Core Need Fulfillment and Interaction Quality predicting Outgroup Attitudes.\\
  General: Fixed effects meta analytic results are presented for completeness only.}
\end{figure}

\begin{figure}
  \caption{Robustness Analyses}
  \label{fig:Robustness}
  \begin{subfigure}{\textwidth}
    \caption{}
    \centering\includegraphics[width=0.65\textwidth]{Figures/forestParametricFERobustContact.png}
  \end{subfigure}
  \begin{subfigure}{\textwidth}
    \caption{}
    \centering\includegraphics[width=0.65\textwidth]{Figures/forestParametricFERobustSDT.png}
  \end{subfigure}
  \caption*{Note: \\
  (a) Need Fulfillment and Intergroup Contact predicting Outgroup Attitudes (full sample).\\
  (b) Core Need Fulfillment predicting Outgroup Attitudes, while controlling for self-determination theory needs (intergroup contact sample).\\
  General: Fixed effects meta analytic results are presented for completeness only.}
\end{figure}

\section{Discussion}
% aims re-iterated
The aim of this project was threefold. We (1) tested the basic predictions of the intergroup contact hypothesis and Allport's optimal conditions in real-life extensive longitudinal data. We also (2) proposed that the fulfillment of situational psychological needs meaningfully predicts positive outgroup attitudes. Finally, we (3) assessed whether the fulfillment of psychological needs (and Allport's conditions) have their effect (partially) through perceived interaction quality.

% Contact hypothesis in ESM data [inconsistent in aggregate but consistent in ML analysis]
When considering the results of the three studies jointly, we found mixed results for the basic intergroup contact hypothesis. More specifically, we observe that seperating between between-subjects and within-subjects effects is crucial. For between-subjects effects, we did not observe a consistent predictor of outgroup attitudes when aggregating people's interaction frequencies and average interaction quality; neither interaction quality, or interaction quality, nor their interaction was a significant predictor across all studies (also see Figure \ref{fig:ContactHypothesis}a). However, once we used multilevel modeling to separate effects into a between-subjects and within-subjects part, we found that when having an interaction with an outgroup member (vs. not having an outgroup interaction),intergroup contact is a strong predictor of outgroup attitudes across all three studies (even when taking other non-outgroup interactions into account; also see Figure \ref{fig:ContactHypothesis}b). While this later results is in line with our expectations, the earlier aggregate findings are surprising because most past cross-sectional studies that have investigated the effect of interaction frequency have also considered an aggregate measurement (albeit not one from actual daily reports). Even though this inconsistency with past research might be a data artefact (e.g., because most people reported substantially more measurements during which they did not have an outgroup interaction), the absence of this aggregate effect could underline the fact that cross-sectional retrospective data might might be misleading because (1) it presents a mixture of within and between-subjects effects \citep[][]{Hamaker2020}, or (2) suffers from recall biases (e.g., where times with no outgroup interactions are undervalued by participants during the retrospective evaluations). In summary, we thus find that when we aggregate interaction frequencies and interaction quality from our intensive longitudinal data we find inconsistent results from the past literature but in line with past literature within participants intergroup contacts are significantly related with positive outgroup attitudes.

@@

% Need fulfillment in intergroup contact [consistent predictor, through quality, slightly better than Allport and SDT]
Finally, when looking at the results for our main proposal regarding the importance of situational key needs, we find highly consistent positive results across all three studies. We find that in all three extensive longitudinal datasets, the fulfillment of core needs during intergroup contacts predicts higher interaction quality perceptions and more positive outgroup attitudes. We find that in all three studies the effect of core need fulfillment on outgroup attitudes goes either fully or partially through interaction quality. We also find that need fulfillment is an important predictor even when taking other fundamental psychological needs or Allport's conditions into account. In fact, our core situational needs measure predicted outgroup attitudes slightly better than Allport's conditions and consistently explained more variance in outgroup attitudes than any of the other psychological needs. In most cases the core need even took over the variances previously explained by the self-determination theory needs. We thus find strong evidence that within everyday life interactions of recent migrants with majority outgroup members, the perception that one's interaction specific psychological needs are fulfilled offers a meaningful and flexible predictor of outgroup attitudes. As such the data offers a first test of the motivational basis of intergroup contact.

\subsection{Limitations}
% (1) Sample = minority and (voluntary) migrants, (2) Methodology = short scales (reliability) and no lagged effects, (3) Core Need Concept = non-specific
As with any approach our studies are not without practical and theoretical limitations. One limitation is our choice of samples. While we believe that a need–based mechanism should be relevant to any inter-group contact, our sample has focused on a minority- and (voluntary) migrant perspective. Without additional evidence it thus remains difficult to judge whether motivational effects will generalize to other migrant groups (e.g., forced migrants), other intergroup contexts (e.g., gender- or sexual orientation groups), and to majority groups in their outgroup attitudes. We know of no research that for other contexts psychological needs would not be relevant (once existential needs are met) but future research may extend our findings to other contexts to build an even broader understanding of psychological needs in intergroup contacts. A second limitation lies in our methodology. While extensive longitudinal data is close to the real-life events this method comes at the expense of longer and more robust scales. Long and repetitive scales are often not feasible in extensive longitudinal methods because of the increased burden to the participants. To circumvent this shortcoming, we have ensured that the measures we used were, whenever possible, based on past validations. However, the circumstance remains that extensive longitudinal data often does not allow the same scrutiny of measurement reliability as single–shot cross–section data sets. An additional methodological question lies in the unexplored potential of the longitudinal aspects of our data. For our research questions we have focused on contemporaneous effects within the data set, yet future investigations should seek to extend the mechanism to developmental trajectories within and between participants. And finally, our conceptualization of core situational needs has been focused on the most essential test of a motivational mechanism. This has come at the expense of specificity in the situational needs (i.e., we have not explored which exact needs people had during their interactions). With the adaptive measurement we have proposed such an investigation would be possible (e.g., by clustering or classifying the free text entries) but would not have been relevant to our theory–focused research question. Nonetheless, future research may explore which exact needs are most important in different intergroup contexts.

\subsection{Implications}
% ESM data
Given these limitations, we can draw a number of implications for other researchers and practitioners from our results. A first implication concerns the feasibility and usefulness of extensive longitudinal data for intergroup contact research and the broader field of social psychology. While setting up an extensive longitudinal study is in no manner easy, given the amount and quality of the data, we believe the efforts to be similar to a sizable cross-section data collection (i.e., for a longitudinal or a high-quality cross-sectional data set with 10,000 data-points). Given the comparable effort, this opens up the possibility to explore research questions that focus on real-life phenomena outside the lab or focus on phenomena that depend on changes and influences over time. In the context of intergroup contact research, we were among the first to answer calls to test intergroup contact mechanisms using on extended real-life data \citep[e.g.,][]{Pettigrew2011, MacInnis2015}. And in doing so, we not only collected a sizable amount of real-life data, but our consideration of extensive longitudinal data may present new inconsistencies in how participants perceive and cognitively aggregate their past interactions with other groups — which may suggest large-scale recall-biases or conflations of within and between participant effects in conventional cross-section studies.

% Motivation and core needs 
A broader theoretical implication relates to role of situational motivation in intergroup contacts. Our results offer a first promising test of psychological needs in intergroup contact. While our results are tentative given their novelty within the field, they were highly consistent across studies and may offer new theoretical avenues moving forward. Psychological needs are a facet of the human experience that has thus far been under–emphasized in the intergroup contact literature. This stands in stark contrast of the many cognitive \citep[e.g.,][]{Pettigrew1998, Brown2005} and emotional aspects investigated within the field \citep[e.g.,][]{Stephan2008, Paolini2004}. However, if motivational mechanisms are indeed shown to be meaningful factors in intergroup contacts, future research may be able to integrate broader theoretical frameworks of intergroup contact (e.g., motivations guiding cognition and affect, which in turn drive behavior. cf., theory of reasoned goal pursuit; \citealp{Ajzen2019}).

% practical use of needs
Additionally, situational motivations in intergroup contact also offer promising avenues for practitioners and policy-makers. Intergroup contact theory is among the most implemented psychological theories \citep[e.g.,][]{Pettigrew2006, AlRamiah2012a, Reimer2021}. Given our findings that psychological needs in every-day intergroup contacts were at least as powerful as Allport's conditions in predicting outgroup attitudes, considerations of people's needs offer a substantially simpler mechanism to address. In cases where some or all optimal contact conditions are not possible to be fulfilled, needs offer an even more compelling alternative (e.g., where equal status is contextually not possible or in cases where people help despite a lack of institutional support). Additionally, our conceptualization of situational needs might offer an additional opportunity for practitioners and interventions. Instead of addressing needs as a one-size fits all solution (e.g., simply focusing on competence needs), one may at times ask outgroup interaction partners what they need during an interaction. This is not to say that we should not explore which needs tend to be relevant to specific groups in specific intergroup contact contexts. Rather, interventions for which such data is not available or infeasible to collect, a flexible and reactive approach to intergroup contact needs might be more fruitful.

% Conclusion paragraph
In sum, in this paper we used extensive longitudinal methodologies to capture real-life data of recent migrants in their interactions with the majority outgroup. Our three studies showcase the feasibility and utility of such data to test intergroup contact theory. We provide evidence that the fulfillment of situational psychological needs during real-life intergroup contacts meaningfully predicts perceived interaction quality and positive outgroup attitudes. Our results point to motivation as an understudied aspect of intergroup contact that is important in understanding when and why an interaction is perceived as positive and will lead to more positive outgroup attitudes.

>>>>

% define document type (i.e., template. Here: A4 APA manuscript with 12pt font)
\documentclass[man, 12pt, a4paper, mask]{apa7}

@@

% Title and header
\title{Psychological Needs During Intergroup Contact: Three Experience Sampling Studies}
\shorttitle{Needs in Intergroup Contact}

% Authors
\author[*,1]{Jannis Kreienkamp}
\author[1]{Maximilian Agostini}
\author[1]{Laura F. Bringmann}
\author[1]{Peter de Jonge}
\author[1]{Kai Epstude}
\affiliation{\hfill}

\affil[1]{University of Groningen, Department of Psychology}

@@

We have no known conflict of interest to declare. The authors received no specific funding for this work. Materials and  software is available at \url{https://janniscodes.github.io/intergroup-contact-needs/}  \citep{KreienkampMasked2022}. Protocols, materials, data, and code are available at \url{https://osf.io/pr9zs/?view_only=1ea47bb646694632a764dead807ef970} \citep{KreienkampMasked2022a}.

@@

% Abstract
\abstract{
One challenge of modern intergroup contact research has been the question of when and why an interaction is perceived as positive and leads to better intergroup relations. This question becomes especially relevant when considering that negative intergroup contacts are common during everyday life. To understand positive intergroup interactions, we propose to consider situational motivations and psychological needs during everyday intergroup contact. We conducted three extensive longitudinal studies with recent migrants, to capture their interactions with the majority outgroup (total N of measurements = 10,297). Across the three studies, we find inconsistent results for a traditional test of the contact hypothesis with experience sampling data. However, when using multilevel models to test the contact hypothesis, we find that individual outgroup interactions and perceived interaction quality have a strong influence on positive outgroup attitudes. We find that the fulfillment of psychological needs offers a strong predictor of positive attitudes after interactions and that these positive effects emerged via perceived interaction quality. The situational needs remained the core predictor of outgroup attitudes even after controlling for Allport’s (\citeyear{Allport1954b}) contact conditions or more general fundamental needs. As one of the first studies to test intergroup contact theory using extensive longitudinal data, we offer insight into the mechanisms of positive intergroup contact during real-life interactions and find situational motivations to be a key building block of understanding and addressing positive intergroup interactions.

\noindent\textbf{Public significance statement}: In this paper, we provide evidence that the fulfillment of situational psychological needs during real-life intergroup contacts meaningfully predicts perceived interaction quality and positive outgroup attitudes. Methodologically, this offers testament to the emerging practice of capturing real-life interactions using extensive longitudinal data. Theoretically, our results give weight to motivational fulfillment as a flexible and relevant mechanism in understanding positive intergroup contact.
}

@@

\begin{figure}
  \caption{Contact Hypothesis}
  \label{fig:ContactHypothesis}
  \centering\includegraphics[width=\textwidth]{Figures/forestParametricGeneralComb.png}
  %\begin{subfigure}{\textwidth}
  %  \caption{}
  %  \centering\includegraphics[width=0.75\textwidth]{Figures/forestParametricGeneralLm.png}
  %\end{subfigure}
  %\begin{subfigure}{\textwidth}
  %  \caption{}
  %  \centering\includegraphics[width=0.75\textwidth]{Figures/forestParametricFEGeneralLmer.png}
  %\end{subfigure}
  \caption*{Note: \\
  (a) summary of regression results from the aggregated contact and interaction quality data.\\
  (b) summary of mixed models results of the contemporaneous contact effects.\\
  General: Fixed effects meta analytic results are presented for completeness only.}
\end{figure}

\begin{figure}
  \caption{Core Need Fulfillment}
  \label{fig:AllportNeedFulfillment}
  \centering\includegraphics[width=\textwidth]{Figures/forestParametricTheoryComb.png}
  %\begin{subfigure}{\textwidth}
  %  \caption{}
  %  \centering\includegraphics[width=0.6\textwidth]{Figures/forestParametricFETheoryQualityCore.png}
  %\end{subfigure}
  %\begin{subfigure}{\textwidth}
  %  \caption{}
  %  \centering\includegraphics[width=0.6\textwidth]{Figures/forestParametricFETheoryAttitudeCore.png}
  %\end{subfigure}
  %\begin{subfigure}{\textwidth}
  %  \caption{}
  %  \centering\includegraphics[width=0.6\textwidth]{Figures/forestParametricFETheoryAttitudeCoreQuality.png}
  %\end{subfigure}
  \caption*{Note: \\
  (a) Core Need Fulfillment predicting Interaction Quality.\\
  (b) Core Need Fulfillment predicting Outgroup Attitudes.\\
  (c) Core Need Fulfillment and Interaction Quality predicting Outgroup Attitudes.\\
  General: Fixed effects meta analytic results are presented for completeness only.}
\end{figure}

\begin{figure}
  \caption{Robustness Analyses}
  \label{fig:Robustness}
  \centering\includegraphics[width=\textwidth]{Figures/forestParametricRobustnessComb.png}
  %\begin{subfigure}{\textwidth}
  %  \caption{}
  %  \centering\includegraphics[width=0.65\textwidth]{Figures/forestParametricFERobustContact.png}
  %\end{subfigure}
  %\begin{subfigure}{\textwidth}
  %  \caption{}
  %  \centering\includegraphics[width=0.65\textwidth]{Figures/forestParametricFERobustSDT.png}
  %\end{subfigure}
  \caption*{Note: \\
  (a) Need Fulfillment and Intergroup Contact predicting Outgroup Attitudes (full sample).\\
  (b) Core Need Fulfillment predicting Outgroup Attitudes, while controlling for self-determination theory needs (intergroup contact sample).\\
  General: Fixed effects meta analytic results are presented for completeness only.}
\end{figure}

\section{Discussion}
% aims re-iterated
The aim of this project was in essence threefold. We (1) tested the basic predictions of the intergroup contact hypothesis and Allport's optimal conditions in real-life extensive longitudinal data. As a flexible alternative, we (2) proposed that the fulfillment of situational psychological needs meaningfully predicts positive outgroup attitudes. Finally, we (3) assessed whether the fulfillment of psychological needs (and Allport's conditions) have their effect (partially) through perceived interaction quality.

% Contact hypothesis in ESM data [inconsistent in aggregate but consistent in ML analysis]
When considering the results of the three studies jointly, we found mixed results for the basic intergroup contact hypothesis. More specifically, we observe that separating between-participant and within-participant effects is crucial. For between-subjects effects (using aggregated reports), neither the number of interactions, nor the average interaction quality were significant predictors of positive outgroup attitudes across all studies (also see Figure \ref{fig:ContactHypothesis}a). The aggregate findings are surprising because most past cross-sectional studies that have investigated the effect of interaction frequency have also considered an aggregate measurement (albeit not one from actual daily reports). The absence of this aggregate effect could underline the fact that cross-sectional retrospective data might be misleading because (1) it presents a mixture of within and between-subjects effects \citep[][]{Hamaker2020}, or (2) suffers from recall biases (e.g., where times with no outgroup interactions are undervalued by participants during the retrospective evaluations).\footnote{It should be noted that such inconsistencies with past research might in part be a data artefact (e.g., because most people reported substantially more measurements during which they did not have an outgroup interaction). However, the multilevel models we employed should in principle be resistant to such imbalances.} In support of this possibility, once we used multilevel modeling to separate effects into a between-subjects and within-subjects part, we found that when having an interaction with an outgroup member (vs. not having an outgroup interaction) is a strong predictor of outgroup attitudes across all three studies (even when taking other non-outgroup interactions into account; also see Figure \ref{fig:ContactHypothesis}b). We thus find that when we aggregate interaction frequencies and interaction quality from our intensive longitudinal data, we find inconsistent results from the past literature, however within participants, intergroup contacts are significantly related with positive outgroup attitudes.

@@

% Need fulfillment in intergroup contact [consistent predictor, through quality, slightly better than Allport and SDT]
Finally, when looking at the results for our main proposal regarding the importance of situational key needs, we find highly consistent positive results across all three studies. We find that in all three extensive longitudinal datasets, the fulfillment of core needs during intergroup contacts predicts higher interaction quality perceptions and more positive outgroup attitudes. We find that in all three studies the effect of core need fulfillment on outgroup attitudes goes either fully or partially through interaction quality. We also find that need fulfillment is an important predictor even when taking other fundamental psychological needs or Allport's conditions into account. In fact, our core situational needs measure predicted outgroup attitudes slightly better than Allport's conditions and consistently explained more variance in outgroup attitudes than any of the other psychological needs. In most cases, the core need even took over the variances previously explained by the self-determination theory needs. We thus find strong evidence that within everyday life interactions of recent migrants with majority outgroup members, the perception that one's interaction specific psychological needs are fulfilled offers a meaningful and flexible predictor of outgroup attitudes.

\subsection{Limitations}
% (1) Sample = minority and (voluntary) migrants, (2) Methodology = short scales (reliability) and no lagged effects, (3) Core Need Concept = non-specific
Among the limitations of our studies we should consider the following. One limitation is our choice of samples. While we believe that a need–based mechanism should be relevant to any inter-group contact, our sample has focused on a minority- and (voluntary) migrant perspective. Without additional evidence it thus remains difficult to judge whether motivational effects will generalize to other migrant groups (e.g., forced migrants), other intergroup contexts (e.g., gender- or sexual orientation groups), and to majority groups in their outgroup attitudes. We know of no research that for other contexts psychological needs would not be relevant (once existential needs are met) but future research may extend our findings to other contexts to build an even broader understanding of psychological needs in intergroup contacts. A second limitation lies in our methodology. While extensive longitudinal data is close to the real-life events, this method comes at the expense of longer and more robust scales. Long and repetitive scales are often not feasible in extensive longitudinal methods because of the increased burden to the participants. To circumvent this shortcoming, we have ensured that the measures we used were, whenever possible, based on past validations. However, the circumstance remains that extensive longitudinal data often does not allow the same scrutiny of measurement reliability as single–shot cross–section data sets. An additional methodological question lies in the unexplored potential of the longitudinal aspects of our data. For our research questions, we have focused on contemporaneous effects within the data set, yet future investigations should seek to extend the mechanism to developmental trajectories within and between participants. And finally, our conceptualization of core situational needs has been focused on the most essential test of a motivational mechanism. This has come at the expense of specificity in the situational needs (i.e., we have not explored which exact needs people had during specific interactions). With the adaptive measurement we have proposed such an investigation would be possible (e.g., by clustering or classifying the free text entries) but would not have been relevant to our theory–focused research question. Nonetheless, future research may explore which exact needs are most important in different intergroup contexts.

\subsection{Implications}
% ESM data
Given these limitations, we can draw a number of implications for other researchers and practitioners from our results. A first implication concerns the feasibility and usefulness of extensive longitudinal data for intergroup contact research and the broader field of social psychology. While setting up an extensive longitudinal study is in no manner easy, given the amount and quality of the data, we believe the efforts to be similar to a sizable cross-section data collection (i.e., for a longitudinal or a high-quality cross-sectional data set with 10,000 data-points). Given the comparable effort, this opens up the possibility to explore research questions that focus on real-life phenomena outside the lab or focus on phenomena that depend on changes and influences over time. In the context of intergroup contact research, we were among the first to answer calls to test intergroup contact mechanisms using on extended real-life data \citep[e.g.,][]{Pettigrew2011, MacInnis2015}. In doing so, we not only collected a sizable amount of real-life data, but our consideration of extensive longitudinal data may present new inconsistencies in how participants perceive and cognitively aggregate their past interactions with other groups — which may suggest large-scale recall-biases or conflations of within and between participant effects in conventional cross-section studies.

% Motivation and core needs 
A broader theoretical implication relates to role of situational motivation in intergroup contacts. Our results offer a first promising test of psychological needs in intergroup contact. While our results are tentative given their novelty within the field, they were highly consistent across studies and may offer new theoretical avenues. Psychological needs are a facet of the human experience that has thus far been under–emphasized in the intergroup contact literature. This stands in stark contrast of the many cognitive \citep[e.g.,][]{Pettigrew1998, Brown2005} and emotional aspects investigated within the field \citep[e.g.,][]{Stephan2008, Paolini2004}. However, if motivational mechanisms are indeed shown to be meaningful factors in intergroup contacts, future research may be able to integrate broader theoretical frameworks of intergroup contact (e.g., motivations guiding cognition and affect, which in turn drive behavior. cf., theory of reasoned goal pursuit; \citealp{Ajzen2019}).

% practical use of needs
Additionally, situational motivations in intergroup contact also offer promising avenues for practitioners and policy-makers. Intergroup contact theory is among the most implemented psychological theories \citep[e.g.,][]{Pettigrew2006, AlRamiah2012a, Reimer2021}. Given our findings that psychological needs in every-day intergroup contacts were at least as powerful as Allport's conditions in predicting outgroup attitudes, considerations of people's needs offer a substantially more immediate mechanism to address. In cases where some or all optimal contact conditions are not possible to be fulfilled, needs offer an even more compelling alternative (e.g., where equal status is contextually not possible or in cases where people help despite a lack of institutional support). Additionally, our conceptualization of situational needs might offer an additional opportunity for practitioners and interventions. Instead of addressing needs as a one-size fits all solution (e.g., simply focusing on competence needs), one may at times ask outgroup interaction partners what they need during an interaction. This is not to say that we should not explore which needs tend to be relevant to specific groups in specific intergroup contact contexts. Rather, interventions for which such data is not available or infeasible to collect, a flexible and reactive approach to intergroup contact needs might be more fruitful.

\subsection{Conclusion}
% Conclusion paragraph
In sum, in this paper we used extensive longitudinal methodologies to capture real-life interactions of recent migrants with the majority outgroup. Our three studies showcase the feasibility and utility of such data to test intergroup contact theory. We provide evidence that the fulfillment of situational psychological needs during real-life intergroup contacts meaningfully predicts perceived interaction quality and positive outgroup attitudes. Our results point to motivation as an understudied aspect of intergroup contact that is important in understanding when and why an interaction is perceived as positive and will lead to more positive outgroup attitudes.
